### 2020-12-29 msccls/MScTemplateCS.tex
2500 ord

\documentclass[nofilelist]{cslthse-msc}
% to show a list of used packages at the end of the document, delete the nofilelist option
%\documentclass{cslthse-msc} 
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{amsmath}
%\usepackage{amsfonts}
%%\usepackage{amssymb}
\usepackage{amsthm}
%\usepackage{makeidx}
\usepackage{graphicx}
\usepackage[titletoc, header, page]{appendix}
\usepackage{transparent}
\usepackage{natbib}
\usepackage{wrapfig}
\usepackage{multirow}
\usepackage{amsmath}

% used to display the used files at the end. Select nofilelist as a package option to disable this
\listfiles % initialize

%\geometry{showframe}
%better like this?
%\student{Flavius Gruian}{Flavius.Gruian@cs.lth.se}
\students{Nick Persson}{dat14npe@student.lu.se}{Filip Karabeleski}{dat14fka.student.lu.se}

\thesisnumber{LU-CS-EX: 2020-XX} % Birger Swahn will provide this number to you, once the thesis is ready for publication
% default is Master. Uncomment the following for "kandidatarbete"/Bachelor's thesis
%\thesistype{Bachelor}{Kandidatarbete}

%\title{Formatting a Master's Thesis}
\title{Keyword Categorization and Emotion Analysis in Voice Calls}

%\onelinetitle
%\twolinestitle
\threelinestitle
%\fourlinestitle

\subtitle{Subtitle WIP}
\company{Telavox AB}
\supervisors
{Sara Lindgren, \href{mailto:Sara.Lindgren@telavox.com}{\texttt{Sara.Lindgren@telavox.com}}}
{Pierre Nugues, \href{mailto:Pierre.Nugues@cs.lth.se}{\texttt{Pierre.Nugues@cs.lth.se.com}}}
\examiner
{Flavius Gruian,\href{mailto:Flavius.Gruian@cs.lth.se}{\texttt{Flavius.Gruian@cs.lth.se}}}

\date{\today}
%\date{January 16, 2015}

\acknowledgements{
If you want to thank people, do it here, on a separate right-hand page. Both the U.S. \textit{acknowledgments} and the British \textit{acknowledgements} spellings are acceptable.

We would like to thank Lennart Andersson for his feedback on this template.

We would also like thank Camilla Lekebjer for her contribution on this template, as well as Magnus Hultin for his popular science summary class and example document.

Thanks also go to the following (former) students for helping with feedback and suggestions on this template: Mikael Persson, Christoffer Lundgren, Mahmoud Nasser.
}

\theabstract{
This document describes the Master's Thesis format for the theses carried out at 
the Department of Computer Science, Lund University.

Your abstract should capture, in English, the whole thesis with focus on the problem and solution in 150 words. It should be placed on a separate right-hand page, with an additional \textit{1cm} margin on both left and right. Avoid acronyms, footnotes, and references in the abstract if possible.

Leave a \textit{2cm} vertical space after the abstract and provide a few keywords relevant for your report. Use five to six words, of which at most two should be from the title.
}

\keywords{Machine Learning, NLP, Sentiment Analysis, Keyword Extraction, (style, structure)}

%% Only used to display font sizes
\makeatletter
\newcommand\thefontsize[1]{{#1 \f@size pt\par}}
\makeatother
%%%%%%%%%%

\begin{document}
\renewcommand{\bibname}{References}

\makefrontmatter
\chapter{Introduction}

\section{Background}
Natural language processing (NLP) is a field of study mainly in computer science that deals with analyzing human language in a way that computers can understand and process. According to \citet{ntlk2009}, natural language is used for everyday communication by humans, in contrast to artificial languages such as programming languages and mathematical notations. Since natural languages continuously develop as time goes on and its adherence to structures diminish, it is difficult to define explicit rules to follow.

\subsection{Emotion Recognition}
Emotion has always been a key part of human language, and it therefore comes naturally that teaching a computer to understand emotion in human language is a big part of NLP. Initially, the focus lied in sentiment analysis, a field of study focused on determining the binary polarization of piece of text as either positive or negative. Godtycklig bit om sentiment analysis. 
Emotion recognition is an extension of sentiment analysis that instead of just determining the polarization of text, attributes emotion to it. This process is more challenging than sentiment analysis, as the differences between emotions are subtler than that of just positive and negative. 
Emotion recognition has also been extended to work on different modalities, such as for example visual and auditory input.

Emotion recognition can be applied to data in various modalities, 
such as visual, in the form of facial expressions and body language, audible and text based in the form of dialogues. To o

The LSTM architecture \citep{ntlk2009} is something....

\begin{itemize}
    \item Förklara vad systemet är
    \item Berätta om företaget Telavox
\end{itemize}

\section{Purpose}
\begin{itemize}
    \item Varför behöver problemet lösas, nytta
\end{itemize}
\subsection{Telavox}
Telavox is a telecommunication company that specializes in business-to-business and business-to-customer unified communication as a service. 
\section{Research questions}
The primary aim of this thesis is to investigate whether or not voice call analysis using NLP can be effective enough to be used as a replacement for manual analysis. \\
\textit{What level of precision can be reached with the categorization of keywords from voice calls?}\\
\textit{How well will our algorithm perform compared to other algorithms carrying out the same measurements?}\\
\textit{What level of effectiveness can be reached compared to a human manually analyzing the data?}

\section{Terminology}
NLP, ML, Speech-to-text och (STT), API. Modality. RNN, GRU, LSTM, BPTT(?)

Kan också kallas för Glossary and Abbreviations
\section{Delimitations}
Delimitations are choices made by the researcher which should be mentioned. They describe the boundaries that you have set for the study.
\section{Contributions}
Potentialen av applications 
\section{Related work}

\chapter{Theory}

blablabla information om de olika fields of study
Deep learning is mostly successive neural networks but there are other types as well. Deep belief networks.

\begin{figure}[h!]
    \centering
    \includegraphics[scale=0.4]{msccls/explanatory_images/map_of_techniques.png}
    \caption{Relations between the related fields of study }
    \label{fig:map}
\end{figure}

\section{Machine Learning}
Machine learning \citep{franoischollet2017learning} is a field of study in artificial intelligence that aims to train a system to solve a problem, rather than to explicitly program it. By being presented with large amounts of data associated with the task at hand, the system algorithmically builds a statistical model that discovers structures in the data and allows the system to identify rules for the problem. 
In \textbf{supervised learning} \citep{100pageBurkov}, which is one of the two main types of tasks in machine learning, the dataset is the collection of labeled examples $\{(\mathbf{x}_i, y_i)\}_{i=1}^N.$ Each element $\mathbf{x}_i$ is a feature vector that contains a value describing the example somehow and $y_i$ is a label, which is either a finite set of classes or a real number. The purpose of supervised learning is to use the dataset to produce a model that takes a feature vector x as input and outputs information that allows deducing the label y for this feature vector.) 
The other main type of tasks in machine learning is \textbf{unsupervised learning}. The dataset for unsupervised learning is $\{ (\mathbf{x}_i)\}_{i=1}^N$ where $\mathbf{x}_i$ is a feature vector. The goal of unsupervised learning is to use the dataset to produce a model that takes a feature vector x as input and outputs either another vector or a value.

\subsection{Neural Networks and Deep Learning}

Neural networks are machine learning systems loosely inspired by biological neural networks. A neural network is a collection of connected nodes or neurons that was first mentioned in \citet{mcculloch1943}. A neuron \citep{dawson1998ann} is an information-processing unit that receives inputs and produces a single output that can be sent to multiple other neurons. To find the output of a neuron the weighted sum of all inputs is weighted by the weights all connecting to the neuron. This weighted sum is then passed through an activation function to provide the output.

\begin{figure}[htp]
    \centering
    \includegraphics[width=12cm]{msccls/explanatory_images/single_neuron.png}
    \caption{An example of a single neuron}
    \label{fig:neuron}
\end{figure}

\begin{equation}
    g(x_1, x_2, x_3,...,x_n) = g(x) = \sum_{i=1}^n  xi
\end{equation}

Here, $g(x)$ takes the inputs $x_1, x_2, x_3,...,x_n$, while the activation function $f(x)$ defines the output. An example of a well known activation function is \textit{ReLU}, Rectified linear unit, which has the following function.

\begin{equation}
y = f(g(x)) =
\begin{cases}
  0, & \text{if}\ x \leq 0 \\
  x, & \text{if}\ x > 0
\end{cases}
\end{equation}

Neural networks can in turn be structured together and form multiple layers that progressively learn. These networks fall under the deep learning umbrella. Deep learning has several closely relevant definitions, where \citet{deng2014deep} mentions the following:

\begin{quote}{A class of machine learning techniques that exploit many layers of non-linear information processing for supervised or unsupervised feature extraction and transformation, and for pattern analysis and classification}
\end{quote}

Each of the layers in a deep learning network manipulate their input data in some way, and the result is stored in the layer's weights. The results are monitored with the help of the \textit{loss function}. The loss function returns a distance score that gives an indication on how far off the networks previous prediction was from the true values \citep{franoischollet2017learning}. The phenomena of reducing the distance score that the loss function produces is called training. 
Using the score produced by the loss function as feedback, the \textit{backpropagation} algorithm computes the \textit{gradient}, a multi variable generalization of the derivative, which is then used to adjust the weights in such a way that the distance score produced by the loss function is minimized.

\subsection{Gated Recurrent Unit - GRU}
A gated recurrent unit (GRU) is neural network architecture introduced by \citet{cho2014learning} and is considered an improvement to the traditional recurrent neural networks. Basic GRU nodes are primarily constructed of an \textit{update gate} and a \textit{reset gate}. The update gate regulates the amount of information that is passed along to the next state, while the reset gate on the other hand decides the amount of past information that is going to be disregarded henceforth. 
Although fairly new it is already a well-established variant of recurrent neural networks. In comparison to LSTM layers, GRU layers are cheaper to run since they use less parameters, but have less representational power. It has been shown that GRU layers have had better performance on smaller and less frequent datasets than LSTM layers \citep{Gruber2020AreGC}.

\begin{figure}[h!]
    \centering
    \includegraphics[scale=0.4]{msccls/explanatory_images/gru.png}
    \caption{A GRU node}
    \label{fig:gru_node}
\end{figure}

\subsection{Long Short-Term Memory - LSTM}
Another widely used neural network architecture is the Long Short-Term Memory (LSTM). It was proposed by \citet{hochreiter1997} as a solution to the vanishing gradient problem \citep{hochreiter1998}, that might occur in recurrent neural networks, and has barely been changed since. The vanishing gradient problem occurs in machine learning when the gradient progressively decreases, which prevents the adjustments of the weight's values.

As illustrated in \ref{fig:lstm_node} a LSTM node consists of an \textit{input gate}, an \textit{output gate}, a \textit{forget gate} and a \textit{cell state}. 
\begin{itemize}
    \item The input gate controls the extent of data to write onto the internal cell-state.
    \item The output gate controls the extent of data to pass as output. 
    \item The forget gate, similarly to the GRUs reset gate, decides how much of the previous data is irrelevant and will be ignored. 
    \item The cell state stores the previous state of a node and passes it on to its following node. 
\end{itemize}

\begin{figure}[h!]
    \centering
    \includegraphics[scale=0.4]{msccls/explanatory_images/lstm.png}
    \caption{A LSTM node}
    \label{fig:lstm_node}
\end{figure}

\section{Natural Language Processing - NLP}
Känns vettigt att ha som section

\section{Word Vectors}
Word vectors are a type of word representation where texts are turned into numbers
\subsection{One-Hot Encoding}
The most basic way of creating a word vector is the one-hot encoding method. This creates a very big and sparse $n \cdot n$ matrix where $n$ is the number of words in the embedding. Each word is then represented by an array filled with zeroes except for one index set to one. For example, the one-hot encoding of the phrase \textit{Why was six afraid of seven} would start with the assignment of values to word:

\begin{center}
    \begin{tabular}{c|c|c|c|c|c}

         Why & was & six & afraid & of & seven \\
         \hline
         1 & 2 & 3 & 4 & 5 & 6 \\
    \end{tabular}
\end{center}

This would then be converted into the following matrix:

\begin{center}
    \begin{tabular}{lllllll}
                            & 1 & 2 & 3 & 4 & 5 & 6 \\ \cline{2-7} 
\multicolumn{1}{l|}{Why}    & 1 & 0 & 0 & 0 & 0 & 0 \\
\multicolumn{1}{l|}{was}    & 0 & 1 & 0 & 0 & 0 & 0 \\
\multicolumn{1}{l|}{six}    & 0 & 0 & 1 & 0 & 0 & 0 \\
\multicolumn{1}{l|}{afraid} & 0 & 0 & 0 & 1 & 0 & 0 \\
\multicolumn{1}{l|}{of}     & 0 & 0 & 0 & 0 & 1 & 0 \\
\multicolumn{1}{l|}{seven}  & 0 & 0 & 0 & 0 & 0 & 1
    \end{tabular}
\end{center}

As mentioned, this method is very basic and creates an unnecessarily large matrix that is largely filled up with nothing. If we wanted to represent a vocabulary of 20 000 words, each word would be represented by 19 999 zeroes and a single one.

The other big issue with it is that no meaning can be extracted from the matrix. If our vocabulary contained the words "cat", "dog" and "hamster", we could easily see that these are all domestic animals but in the space of the one hot vector, "cat" is as close to "dog" as it would be to "airplane".

\subsection{Word Embedding}
To resolve the two aforementioned problems with one-hot encodings, word embeddings were created, also known as dense word vectors. \citep{neuralnetworkmethods} 
Here, instead of creating a sparse, hardcoded vector with equal distance between words, they are learned from data and packed into a lower-dimensional vector where each dimension contains some relevant data about the word. According to \citet{franoischollet2017learning}, word embeddings that are 256-dimensional, 512-dimensional or 1024-dimensional are common when dealing with very large vocabularies, in contrast to the 20 000-dimensional one that would be created if one were to use one-hot encoding. Additionally, the geometric relationship between words can and should reflect the semantic relationship between those same words. This means that synonyms and words often used together are embedded closer to each other, while words seldom used together are more distant in the embedding. In addition to distance, direction could also be a useful tool to extract meaning from the embedding. %Temp for formatting, don't forget it's here

\begin{wrapfigure}{r}{0.25\textwidth}
\includegraphics[width=0.9\linewidth]{msccls/explanatory_images/embedding_direction.png} 
\caption{Small example of a word embedding space}
\label{fig:direction}
\end{wrapfigure}

In figure \ref{fig:direction}, the four words \textit{cat}, \textit{dog}, \textit{wolf} and \textit{tiger} are represented on a geometric plane. We can clearly see some of the possible geometric transformations possible by using direction as a tool. For instance, the same vector that takes us from \textit{dog} to \textit{wolf} also takes us from \textit{cat} to \textit{tiger}. This could be interpreted this vector as "converting a domesticated animal to its wildlife equivalent". Another vector will take us both from \textit{cat} to \textit{dog} and from \textit{tiger} to \textit{wolf} which could be interpreted as "converting a feline to a canine". Two meaningful transformation vectors that are common in the real world are "gender" and "plural" vectors. By applying a "female" vector to \textit{king}, \textit{queen} is obtained and the same vector returns \textit{actress} if instead used on \textit{actor}. "Plural", as should seem obvious by now, if used on \textit{actor} will return \textit{actors}. Typically there are thousands of these types of interpretable direction vectors in a large word embedding.

This bit is after the picture, that should be to the right of the last paragraph

\subsection{RoBERTa}

\section{Evaluation}
There are many different metrics that can be used to evaluate a machine learning model, based on what kind of model and what its purpose is. For classification models, many of the metrics are based on a confusion matrix \citep{FAWCETT2006861}. In a confusion matrix, one of the axis represent the actual class of the instance the model is trying to predict and the other represents what instance the model predicted. The most basic and common confusion matrix is a binary $2 \cdot 2$ matrix, as seen in Figure \ref{fig:confusion}. In binary cases, the instances usually can be represented as a yes or no question, such as \textit{is there a bird in this picture?}

\begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth/2]{msccls/explanatory_images/confusion_matrix.png}
    \caption{Binary Confusion Matrix}
    \label{fig:confusion}
\end{figure}

\begin{itemize}
    \item TP means that the model was correct in predicting positive
    \item TN means that the model was correct in predicting negative
    \item FP means that the model was incorrect in predicting positive
    \item FN means that the model was incorrect in predicting negative
\end{itemize}

Whenever our model makes a prediction on an instance during training, we increment the corresponding cell in the matrix, letting us see and draw conclusions based on how our model performed. 
For cases where we have multi-class classification instead of binary, such as when determining emotions, the same logic and matrix is used but on a bigger scale and with the specific classes instead of positive and negative. The equations provided in subsequent sections to display how to calculate metrics are identical for multi-class context, but the classification instead represents the sum of the classification for each class. For instance, if determining if an emotion is \textit{Neutral}, \textit{Happy}, or \textit{Angry}, $TP_{n}$ would represent when it got neutral correct, and TP would be the sum of all emotion predictions it got correct.

%kan snacka om Type I och Type II errors om mer text behövs.

\subsection{Accuracy}
One of the simplest metrics used, accuracy quantifies how many percent of predictions that the model got right and is calculated by taking all instances where our model predicted positive and dividing it with all instances.

$$\frac{TP + TN}{TP + TN + FP + FN}$$

If we have a balanced data set with a similar amount of positive and negative instances, this can give valuable information on the performance of the model since getting predictions right is the goal of any model. If, however, our set is imbalanced and contains e.g. 90 percent negative instances and only 10 percent positive, the model can achieve a 90 percent accuracy by simply predicting negative on every instance so the accuracy suddenly becomes useless.

\subsection{Precision}
Precision is a metric that quantifies how many of the instances predicted positive that actually were true. It's calculated by taking all of the instances where the model predicted positive and was correct, and dividing it by all the instances where it predicted positive.

$$\frac{TP}{TP + FP}$$

Precision gives us information on how accurate the model is when predicting an instance to be true. This is useful when dealing with instances where it's important to not give a false positive. If we have a model that is trying to predict whether or not a suspect is guilty of a certain crime, a high precision would make it unlikely for a innocent person to be determined a criminal. 
Precision does not give any information on the number of false negatives present, meaning that a high precision could mean that the model seldom identifies a person, but when it does identify someone, it's likely that said person is guilty.

When looking at multiple classes, the formula remains the same but with TP and FP representing the sum of all TPs and FPs

\subsection{Recall}

How many did we get right, of all that were actually right

False negative, such as illness
\subsection{F1-score}

\subsection{Loss}

\section{Overfitting}

\section{Baseline - COSMIC}
\begin{itemize}
    \item Dataset till input måste följa Googles standard som de följer vid speech to text. 
\end{itemize}

\section{Theory}
NLP technique, categorization, ML, supervised/unsupervised,  
Mechanical turks to 
Finns både text och audio på MELD, \citep{zhang2019modeling} visar att text för sig är bättre än ljud men att text+ljud > text
train, dev(validation), test, man testar på dev under development, först på test efter man är klar

\chapter{Implementation}

\section{Datasets}
Filip
\subsection{Approach} 
\section{Method}
Google Colab
Word level one hot encoding, kanske relevant, avgör beroende på hur mycket mer vi behöver skriva.
\section{Implementation}
Overfitting \\
Data Augmentation \\
importance of finding the correct data-sets \\
\section{Bibtesting, will be removed}
\citep{emotionlinesdataset} \\
\citep{franoischollet2017learning}\\
\citep{beaver2020towards}
\chapter{Speech-to-text}
Vi antar att det är 2 speakers i samtalet, nämn för- och nackdelar. Ta med i limitations
Vi har endast stöd för engelska pga limits i datamängd, dokumentation. 
Google diarization är i Beta.
Använda kategoriserade ord i Googles STTs speechContexts och därefter använda ML igen för att få högre precision på kategorisering. 
Sample rates 8000 vs 48000 t.ex.

Cased eller uncased? Hur fungerar google speech-to-text med cased/uncased?

\chapter{Keyword Extraction}
\chapter{Emotion Analysis}
6 eller 4 känslor
Sentiment analysis, varför valde vi inte det?

\chapter{Evaluation}

\section{Results}

These are our results

% Please add the following required packages to your document preamble:
% \usepackage{multirow}
\begin{center}
\begin{tabular}{|c|cc|cc|cc|cc|}
\hline
\multirow{2}{*}{Methods} & \multicolumn{2}{c|}{MELD 34}            & \multicolumn{2}{c|}{DailyDialog 45}     & \multicolumn{2}{c|}{IEMOCAP 34}         & \multicolumn{2}{c|}{EmoryNLP 23}        \\ \cline{2-9} 
                         & \multicolumn{1}{c|}{W-Avg F1} & Loss & \multicolumn{1}{c|}{W-Avg F1} & Loss & \multicolumn{1}{c|}{W-Avg F1} & Loss & \multicolumn{1}{c|}{W-Avg F1} & Loss \\ \hline
Cosmic                   & 64.14                         & 2.68 & 60.01                         & 2.94 & 58.77                         & 0.96 & 42.55                         & 2.48 \\ \hline
Vår modell               &                               &      &                               &      &                               &      &                               &      \\ \hline
\end{tabular}
\end{center}

\section{Discussion}
This is the discussion about our results

\chapter{Conclusions}

% Should use consistent formatting when it comes to Names ("FirstName LastName", or "F. LastName")
%\printbibliography
\makebibliography{MyMSc}

\begin{appendices}
\chapter{Dummy Appendix}

% display used packages information unless nofilelist is used in the cslthse-msc package option
\printfilelist

%make sure we're on even page with the pop-sci
\checkoddpage
\ifoddpage
\else
   \newpage
   \thispagestyle{empty}
   \mbox{ }
\fi
\includepdf[pages={1}]{popsci/popsci.pdf}
\end{appendices}

\end{document}